Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
add examples for explanation

\paragraph*{Syntax}
Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
structured. Symmetric DNF lenses, which provide lens functions between DNF
regular expressions, consist of a list of symmetric sequence lenses, as well as
some additional information. Symmetric sequence lenses, which provide lens
functions betwen sequences, consist of a list of symmetric atom lenses, as well
as some additional information. Symmetric atom lenses, which provide lens
functions between atoms, are merely iterations of symmetric DNF lenses. In the
following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
strings.
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{}l@{}}
    % REGEX
    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
                          \ldots\SeqLSep
                          (i_p,j_p,\SAtomLens_p)}
                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
                           \ldots\DNFLSep
                           (i_p,j_p,\SSQLens_p)}
                           ,\ListOf{c_1;\ldots;c_q}
                           ,\ListOf{d_1;\ldots;d_r})$ \\
  \end{tabular}
\end{center}

In greater detail, a symmetric DNF Lens \SDNFLens{} consists of three
components. First is the mapping component. The mapping component is a list of
triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between the two
sequences specified by $i$ and $j$. The second component is a list of integers
$c$. The integer, $c_i$ at position $i$ states that when performing a $\CreateR$
on a string matching sequence $\Sequence_i$, it will use the sequence lens that
maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third component is a
list of integers $d$, and follows the same pattern as the second component, but
maps from right to left.

A Sequence lens \SSQLens{} consists of three components. First is the mapping
component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
sequences, respectively. The third component of the triple, \SAtomLens{}, is an atom lens that
maps between the two atoms specified by $i$ and $j$. The second component is a
list of strings, $\String$. The string, $s_i$ at position $i$ states that when
performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
two things. If $\Atom_i$ is mapped to in the mapping component, then performing
the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
not mapped to, then $\String_i$ will be created as a representative of
$\Atom_i$.

While DNF lenses are useful because they permit an efficient greedy synthesis
procedure, the particulars of their typing and semantics are not the primary
intellectual contributions to our work. The details of the typing and semantics
are included in the supplementary material present in the appendix.

\paragraph*{Information-Theoretic Cost Metric}
Instead of searching for \emph{any} term that satisfies the specification, we
search for the \emph{best} term that satisfies the specification. This begs the
question: what metric should we use to determine which term is best?
Intuitively, we aim to synthesize the lens which is ``closest'' to being a
bijection. Somewhat more formally, we aim to synthesize a lens which minimizes
the expected number of bits which must be used to recover the source from the
target, and the target from the source.

Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses.  Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms.  $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented Boomerang's asymmetric lens combinators \sam{all of Boomerang's combinators?} using the symmetric lens combinators.

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
  benchmarks, we not only provide functions that clean the data, but we also
  provide other synchronization functions: if the cleaned data is updated, those
  changes are reflected in the original format.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad-hoc data present in Linux configuration files into a structured, dictionary
  representation. \sam{What does data cleaning mean?}
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

>>>>

\paragraph*{Syntax}
Roughly speaking, symmetric DNF lenses are an n-ary union of symmetric sequence
lenses, which are, in turn, an n-ary sequence of atomic lenses. Atomic lenses
are iterations of symmetric DNF lenses. In the following syntax, the
nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all represent natural
numbers. The nonterminals $\String$ and $\StringAlt$ represent strings.
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{}l@{}}
    % REGEX
    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
                          \ldots\SeqLSep
                          (i_p,j_p,\SAtomLens_p)}
                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
                          ,\ListOf{(\StringAlt_1,\AtomAlt_1);\ldots;(\StringAlt_r,\AtomAlt_r)})$ \\
    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
                           \ldots\DNFLSep
                           (i_p,j_p,\SSQLens_p)}
                           ,\ListOf{c_1;\ldots;c_q}
                           ,\ListOf{d_1;\ldots;d_r})$ \\
  \end{tabular}
\end{center}

As an example, consider the symmetric DNF lens:
$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
is used, and when performing a \CreateL on a string in $\Sequence_2$,
$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
of three components. First is the mapping component. The mapping component is a
list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between
the two sequences specified by $i$ and $j$. The second component is a list of
integers $c$. The integer, $c_i$ at position $i$ states that when performing a
$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
component is a list of integers $d$, and follows the same pattern as the second
component, but maps from right to left.

As an example, consider the symmetric sequence lens:
$(\SSQLensOf{(1,2,\SAtomLens_1)} ,\ListOf{}
,\ListOf{(\StringAlt_1',\AtomAlt_1)})$, synchronizing data between
$\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \String_1}$ and
$\SequenceOf{\StringAlt_0 \SeqSep \AtomAlt_1 \SeqSep \StringAlt_1 \SeqSep
  \AtomAlt_2 \SeqSep \StringAlt_2}$. This lens maps $\Sequence_1$ to
$\SequenceAlt_2$, and leaves $\AtomAlt_1$ disconnected. On a \CreateR, the
string created is a concatenation of $\StringAlt_1'$ with the default from
$\SAtomLens_1$. More generally, a sequence lens \SSQLens{} consists of three
components. First is the mapping component. The mapping component is a list of
triples $(i,j,\SAtomLens)$. In this, $i$ and $j$ refer to atom $i$ and atom $j$
on the left and right sequences, respectively. The third component of the
triple, \SAtomLens{}, is an atom lens that maps between the two atoms specified
by $i$ and $j$. The second component is a list of strings, $\String$. The
string, $s_i$ at position $i$ states that when performing a $\CreateL$ on a
string matching atom $\Atom_i$, it will do one of two things. If $\Atom_i$ is
mapped to in the mapping component, then performing the atom's $\CreateL$ will
actually $\PutL$ into that string. If $\Atom_i$ is not mapped to, then
$\String_i$ will be created as a representative of $\Atom_i$.

While DNF lenses are useful because they permit an efficient greedy synthesis
procedure, the particulars of their typing and semantics are not the primary
intellectual contributions of our work. The details of the typing and semantics
are included in the supplementary material present in the appendix.

\paragraph*{Information-Theoretic Cost Metric}
Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses. Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms. $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

@@

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all Boomerang's asymmetric lens combinators using a combination of
the symmetric lens combinators, as well as other symmetric lens combinators for
the asymmetric theory extensions present in Boomerang, like matching lenses and
quotient lenses.

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
  benchmarks, we not only provide functions that clean the data, but we also
  provide other synchronization functions: if the cleaned data is updated, those
  changes are reflected in the original format.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad-hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}
